Linear Regression
========================================================



```{r soluData}
library(AppliedPredictiveModeling)
data(solubility)
ls(pattern="solT")
dim(solTrainXtrans)
trainSol <- solTrainXtrans
trainSol$Solubility <- solTrainY
```
Lets start modeling
```{r linreg}
lm.all <- lm(Solubility ~ ., data=trainSol)
summary(lm.all)
lmPred.all <- predict(lm.all, newdata=solTestXtrans)
lmValues.all <- data.frame(obs=solTestY, pred=lmPred.all)
defaultSummary(lmValues.all)
```
Robust linear regression,
```{r robustreg}
library(MASS)
rlm.all <- rlm(Solubility ~ ., data=trainSol)
summary(rlm.all)
rlmPred.all <- predict(rlm.all, newdata=solTestXtrans)
rlmValues.all <- data.frame(obs=solTestY, pred=rlmPred.all)
defaultSummary(rlmValues.all)
```
Doesn't do any better

Cross validation
```{r crossvalidate}
ctrl <- trainControl(method="cv", number=10)
set.seed(100)
lm.cv.all <- train(x=solTrainXtrans, y=solTrainY, method="lm", trControl=ctrl)
lm.cv.all
```
Plot results
```{r plots}
xyplot(solTrainY ~ predict(lm.cv.all), 
       type=c("p", "g"),
       xlab="Predicted",
       ylab="Observed")

xyplot(resid(lm.cv.all) ~ predict(lm.cv.all),
       type=c("p", "g"),
       xlab="Predicted",
       ylab="Residuals")
```
Make a smaller model by removing correlated variables
```{r remCor}
corThresh <- 0.75
tooHigh <- findCorrelation(cor(solTrainXtrans), corThresh)
corrPred <- names(solTrainXtrans)[tooHigh]
trainXfltrd <- solTrainXtrans[ , -tooHigh]
testXfltrd <- solTestXtrans[ , -tooHigh]
set.seed(100)
lm.cv.fltrd <- train(x=trainXfltrd, y=solTrainY, method="lm", trControl=ctrl)
lm.cv.fltrd
```
We can also use rlm, but that model does not tolerate singular matrices, so we will have to PCA the data 
first
```{r trainRlm}
set.seed(100)
lm.cv.rlm <- train(solTrainXtrans, solTrainY,
                   method="rlm",
                   preProcess="pca",
                   trControl=ctrl)
lm.cv.rlm
```

Lets compare the all and fltrd models over the test set
```{r compareAllFltrdRlm}
pred.cv.all <- predict(lm.cv.all, newdata=solTestXtrans)
pred.cv.fltrd <- predict(lm.cv.fltrd, newdata=testXfltrd)
pred.cv.rlm <- predict(lm.cv.rlm, newdata=solTestXtrans)
defaultSummary(data.frame(obs=solTestY, pred=pred.cv.all))
defaultSummary(data.frame(obs=solTestY, pred=pred.cv.fltrd))
defaultSummary(data.frame(obs=solTestY, pred=pred.cv.rlm))
```



##Partial Least Squares

```{r pls}
library(pls)
plsFit <- plsr(Solubility ~ ., data=trainSol)
perf.pls <- do.call("rbind", lapply(1:plsFit$ncomp, function(n){
                      pred.pls <- as.vector(predict(plsFit, newdata=solTestXtrans, ncomp=n))
                      defaultSummary(data.frame(obs=solTestY, pred=pred.pls))
                  }))

Ridge-regression
```{r ridge}
library(elasticnet)
ridgeModel <- enet(x=as.matrix(solTrainXtrans), y=solTrainY, lambda=0.001)
