---
title: "Measuring Performance in Regression Models"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document:
    highlight: tango
    number_sections: yes
    theme: journal
author: "Vishal Sood"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE)
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(include=TRUE)
knitr::opts_chunk$set(results="asis")
knitr::opts_chunk$set(fig.width=12)
knitr::opts_chunk$set(fig.height=14)
prjpath <- "~/work/learn/machineLearn/appliedPredictiveModeling/"
datapath <- paste(prjpath, "data/", sep="")
analpath <- paste(prjpath, "analyses/", sep="")
rcodepath <- paste(analpath, "Rcode/", sep="")
setwd(analpath)

```{r loadLibsAndSource, include=FALSE}
reqdpkgs <- c("lattice", "latticeExtra", "ggplot2", "reshape2",
						 	"plyr", "Hmisc", "corrplot", "AppliedPredictiveModeling",
							"caret", "rms", "e1071", "ipred", "MASS")
lapply(reqdpkgs, library, character.only=TRUE)
```

#Quantitative Measures of Performance
RMSE.

$R^2$: proportion of the information in the data that is explained by the model. It is a measure of correlation, not accuracy. Depends on the variation of the outcome.

Spearmans' rank correlation

#The Variance-Bias Trade-off
Bias: reflects how close the functional form of the model can get to the true relationship between the predictors and the outcome.  Low bias means that the model can get very close to the data.

Low variance would be when the variance would not substantially change if another set of data points were generated the same way.

Highly correlated predictors can lead to collinearity issues, which can greatly increase the model variance.
