---
title: "Nonlinear Regression Models"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document:
    highlight: tango
    number_sections: yes
    theme: journal
author: "Vishal Sood"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE)
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(include=TRUE)
knitr::opts_chunk$set(results="asis")
knitr::opts_chunk$set(fig.width=12)
knitr::opts_chunk$set(fig.height=14)
prjpath <- "~/work/learn/machineLearn/appliedPredictiveModeling/"
datapath <- paste(prjpath, "data/", sep="")
analpath <- paste(prjpath, "analyses/", sep="")
rcodepath <- paste(analpath, "Rcode/", sep="")
setwd(analpath)

```{r loadLibsAndSource, include=FALSE}
reqdpkgs <- c("lattice", "latticeExtra", "ggplot2", "reshape2",
						 	"plyr", "Hmisc", "corrplot", "AppliedPredictiveModeling",
							"caret", "rms", "e1071", "ipred", "MASS", "lars", "pls", "stats")
lapply(reqdpkgs, library, character.only=TRUE)
```


#Introduction
Neural networks, mutivariate adaptive regression splines (MARS), support vector machines (SVMs), and $K$-nearest neighbors (KNNs). Next chapter tree-based models.

#Neural Networks
